Evolution of Processes in Delivering Reliable Software in the European Scientific Arena

### 2018-06-15 paper-ieee-trans.tex
fix minor typos and nesting

The SQA criteria are a set of conventions and recommendations that pave the way for
the adequate development, timely delivery and reliable operation of the produced software components.
It emphasizes the quality requirements and best practices to be applied at the
development phase, such as style and testing compliance or human code reviews, to protect the
production source code versions. Furthermore, the criteria promote the adoption of the software
by providing minimum requirements of documentation content and stressing the usage of automated
solutions for software deployment. The SQA criteria is continuously evolving \cite{sqa-baseline}, aiming at serving
as a reference for future software development projects within the European research ecosystem.

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities, are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO--1) and second (INDIGO--2) releases. By INDIGO--2, 53\% of the software components
(17 out of 32) were over the threshold of the aforementioned project's SQA code coverage recommendation
(70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage. 
The indiviual records followed also a growing trend (only 6 out of 32 components had lower values in 
INDIGO--2). It is worth noting that Fig.~\ref{fig:fig_unittest} not only includes new software developments 
implemented from scratch within the project. Software tools and libraries from 
external open--source projects -- such as {\sl tosca-parser} or {\sl heat-translator} --, contributed upstream
by the project, and well--established products, involved in {\sl INDIGO--DataCloud}
but not contributing with the 100\% of their codebases -- such as {\sl dCache} --, are also considered in the
analysis. However the values given apply to the entire codebase, even for the latter case of products. These 
products challenged the application of the SQA policies described in this section, by two means:
\begin{itemize}
    \item Previously developed code was not refactored since it was not \textit{owned} by the project.
    \item An agreement on prevailing SQA policies for such cases where quality practices were already in place.
\end{itemize}

@@

\bibitem{sqa-baseline}
Pablo Orviz et al.,
A set of common software quality assurance baseline criteria for research projects,
2017.
[Online] Available: \url{http://digital.csic.es/bitstream/10261/160086/4/CommonSQA-v2.pdf}

>>>>

The SQA criteria are a set of conventions and recommendations that pave the way for
the adequate development, timely deliver and reliable operation of the produced software components.
It emphasizes the quality requirements and best practices to be applied at the
development phase, such as style and testing compliance or human code reviews, to protect the
production source code versions. Furthermore, the criteria promote the adoption of the software
by providing minimum requirements of documentation content and stressing the usage of automated
solutions for software deployment.

@@

\item \textit{Unit and Functional testing}.
Changes involving the addition of new functionalities, are required to be tested. \textit{Regression
testing} in this context is accomplished by enforcing the definition and periodic execution
of the tests, covering past fixed issues, to ensure that they are not reintroduced during the development
activities. \textit{Unit testing} completes the source code testing evaluation by focusing on the code's
internal design. The SQA criteria of {\sl INDIGO--DataCloud} set a recommended threshold of 70\% coverage
for unit testing. In Fig.~\ref{fig:fig_unittest}, unit testing coverage is compared for each software
component between the two major releases, denoting an incremental trend over the course of the project:
code coverage increased from an average value of 50.32\% to 63.34\%, outlined by the respective dates of
the project's first (INDIGO-1) and second (INDIGO-2) releases. By INDIGO-2, 53\% of the software components
(17 out of 32) were over the threshold of the aforementioned project's SQA code coverage recommendation
(70\%), while 75\% (24 out of 32) of the product stack exceeded 50\% coverage.

@@
